---
title: "Downloading data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{downloading}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A primary use case for the `{AquaMatchr}` package is to facilitate downloading of  [AquaMatch](https://aquasat.github.io/AquaMatch_documentation/), [lakeSR and siteSR](https://aquasat.github.io/AquaMatch_lakeSR/), and [riverSR](https://zenodo.org/records/4304567) data products. These datasets are hosted in different locations and so the functions provided within `{AquaMatchr}` to retrieve them vary. This vignette walks through how `{AquaMatchr}` functions are used to do this. Note that you are free to do this manually by visiting the data repositories directly, but the functions in this package are designed to make the process simpler to do entirely within R.

We start by loading the package and any other packages we'll want to use for data wrangling, etc.

```{r setup, message = FALSE, warning = FALSE}
library(AquaMatchr)
library(dplyr)
library(arrow)
```

# Downloading and saving *in-situ* parameter data

The `AquaMatchr::download_parameters()` function is the method that we've provided to download harmonized *in-situ* sampling data from AquaMatch. It will download the data from the [Environmental Data Initiative ("EDI")](https://edirepository.org/) repository and load it into the R environment as a component item of a `list`. (It does **not** save permanent local copies of the datasets automatically.)

For example, the following code snippet can be used to download the most recent revision of the total suspended solids (TSS) data product:
```{r download-recent-tss}
# By default, this function uses "newest" for version but can be set to an
# integer indicating the version number (1, 2, etc.)
tss_data <- download_parameters(parameters = "tss", version = "newest")

names(tss_data)
```

The data are returned as an item within a list, and if multiple parameters are
requested then they'll each be individual list items. Below we'll request both
TSS and Secchi disk depth (SDD) products:
```{r download-recent-tss-sdd}
# Whatever value is provided to "version" will be used for all parameters that
# are queried
tss_sdd_data <- download_parameters(
  parameters = c("tss", "sdd"),
  version = "newest"
)

# One list item for each parameter requested
names(tss_sdd_data)
```

Note that the datasets are large:
```{r check-param-size}
# TSS size
format(object.size(tss_sdd_data$tss), "MB")

# SDD size
format(object.size(tss_sdd_data$sdd), "MB")
```

If you would like to pull out the data frames from the list to make using them easier, you can do it like this:
```{r deconstruct-list}
tss_df <- tss_sdd_data$tss

sdd_df <- tss_sdd_data$sdd

# Remove the list to save memory
rm(tss_sdd_data)
gc()
```

And finally, if you'd like to save the datasets permanently for use after this R session, you might do it like this (but include a usable file path):
```{r save-param-data, eval = FALSE}
write_feather(x = tss_df, path = "path/to/folder/of/choice")
```

Feather files are both time- and space-efficient ways to store data for use in R and other programming languages. (See, for example, [sections 5.4.2 and 5.4.3](https://csgillespie.github.io/efficientR/input-output.html) in Gillespie & Lovelace, 2021). You could also use a `.csv` or `.rds` format here, for example. We suggest using `.feather` or `.csv` for analyses that will use this package because these are the most common filetypes our functions are built to work with.

## Selecting specific parameter versions

There may be instances where you do not want to use the newest version of a published AquaMatch dataset. You can provide an integer number to the `version` argument in place of "newest" to indicate which version, or "revision", you want to retrieve. If you're not sure, you can check how many revisions exist for each dataset on the EDI page for each data release under *Package ID* \> *all revisions*. For example, for TSS that would bring you to [this page](https://portal.edirepository.org/nis/revisionbrowse?scope=edi&identifier=2048). Doing so currently requests that version number for all parameters in your request, so if you don't want, e.g., version 2 of each dataset, you may wish to do your requests separately. For example:
```{r demonstrate-versions, error = TRUE}
# This will fail because SDD does not have a version 2
download_parameters(
  parameters = c("tss", "sdd"),
  version = 2
)

tss_v2 <- download_parameters(
  parameters = "tss",
  version = 2
)

sdd_v1 <- download_parameters(
  parameters = "sdd",
  version = 1
)
```


# Downloading lakeSR and siteSR

[LakeSR](https://portal.edirepository.org/nis/mapbrowse?packageid=edi.2114.1) and [siteSR](https://portal.edirepository.org/nis/mapbrowse?packageid=edi.2254.1) are housed on EDI, similarly to the harmonized *in-situ* sampling datasets. A separate function is provided to download each one: `AquaMatchr::download_siteSR()` and `AquaMatchr::download_lakeSR()`. Both functions download multiple large files to a path provided by the user. The files are not read into the R environment during the download in order to save memory. Both have `algal_mask` arguments: If `TRUE`, the algal mask version of the dataset (DSWE1a) is downloaded, otherwise DSWE1 is used (i.e., `FALSE`).

For example, the code below would download the DSWE1 version of siteSR and would also ask the user to confirm that any existing files with identical names in the download location should be overwritten (i.e., `ask` = `TRUE`).

```{r sitesr-download, eval = FALSE}
download_siteSR(
  save_location = "path/to/folder/of/choice",
  algal_mask = FALSE,
  ask = TRUE
)
```

The `download_siteSR()` function downloads the following files from EDI:

+ lakeSR handoff coefficients (.csv)
+ siteSR sites list (.csv)
+ siteSR data from Landsat 4 (.feather)
+ siteSR data from Landsat 5 (.feather)
+ siteSR data from Landsat 7 (.feather)
+ siteSR data from Landsat 8 (.feather)
+ siteSR data from Landsat 9 (.feather)

The lakeSR handoff coefficients file is included with this download because both datasets use the same file, but it is only contained in the lakeSR release.

The equivalent `download_lakeSR()` call would download lakeSR files:
```{r lakesr-download, eval = FALSE}
download_lakeSR(
  save_location = "path/to/folder/of/choice",
  algal_mask = FALSE,
  ask = TRUE
)
```

The lakeSR files:

+ lakeSR handoff coefficients (.csv)
+ lakeSR sites list (.csv)
+ lakeSR data from Landsat 4 (.feather)
+ lakeSR data from Landsat 5 (.feather)
+ lakeSR data from Landsat 7 (.feather)
+ lakeSR data from Landsat 8 (.feather)
+ lakeSR data from Landsat 9 (.feather)


# Downloading riverSR

The [riverSR dataset](https://doi.org/10.5281/zenodo.4304567) is stored on Zenodo, and due to its large size we download it without bringing it into the R environment. This takes place through the `AquaMatchr::download_riverSR()` function, which requires a download path to be provided by the user. The `timeout_length` for `download_riverSR()` can also be specified but by default is 4,000 seconds, which typically worked in testing for this function (though it may vary throughout time). RiverSR will be downloaded and saved as `"riverSR_usa_v1.1.feather"` in the folder specified in the `save_location` argument.

```{r riversr-download, eval = FALSE}
download_riverSR(save_location = "path/to/folder/of/choice")

# Can be loaded into environment using
riverSR <- read_feather("path/to/folder/of/choice/riverSR_usa_v1.1.feather")
```

The riverSR download will be nearly 14 GB in size and the download may take 30 minutes or more to complete.




