---
title: "Example AquaMatch workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example AquaMatch workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The goal of this vignette is to demonstrate a full workflow for the `AquaMatchr` package. We demonstrate the code to download harmonized Water Quality Portal (WQP) and siteSR files, stack the siteSR files into a single product, perform 5-day matchups between WQP and siteSR data, and finally apply Gardner polynomial intermission handoffs to the matchup product.

Note that this vignette is intended as a usable template, but by default the R Markdown file is set to `eval = FALSE`.

Lastly, for more detailed information on the steps in this workflow we suggest reviewing the other vignettes in this package. This one is intended more as an example and less as documentation.

## 1. Setup and Directory Prep

For this tutorial, we will save files downloaded and generated by the script to a temporary directory.

> **Note for real-world use:** When running this on your own machine, you should replace `temp_dir` with a real path on your computer (e.g., `"C:/Users/MyName/Documents/aquamatch_example"`) so your data is saved permanently.

```{r setup}
library(AquaMatchr)
library(arrow)
library(dplyr)

# Define our storage directory for this session
temp_dir <- tempdir()
```

## 2. Downloading the Data

*(Note: Depending on your internet connection, these steps may take a few minutes).*

### Chlorophyll *a* data

First, we use `download_parameters()` to fetch the required harmonized WQP chlorophyll *a* data from the Environmental Data Initiative (EDI).

```{r download-chla}
# Defaults to newest version if multiple are available
chla_data <- download_parameters(
  parameters = "chla"
)

# Data are returned as a named list, so we'll extract the data frame from this
# and save it to disk. You could choose to keep it loaded in R, but for the 
# purposes of this example it's not necessary. We use .feather here but that's
# just preference, not a requirement.
write_feather(
  x = chla_data$chla,
  sink = file.path(temp_dir, "chla_harmonized.feather")
)

# Free up memory
rm(chla_data)
gc()
```

### siteSR data

Next we download the siteSR data from EDI using `download_siteSR()`.

```{r download-siteSR}
# Naming this _paths because the function returns a vector of filepaths after it
# completes
siteSR_paths <- download_siteSR(
  save_location = temp_dir,
  # DSWE1
  algal_mask = FALSE,
  # Don't ask about overwriting existing files
  ask = FALSE
)
```

## 3. Stack the siteSR dataset

The siteSR dataset is downloaded as a separate file for each of the Landsat missions it contains. Here we stack them into a single Arrow Table and save it locally.

```{r build-siteSR}
full_sr <- build_sr(
  which_sr = "siteSR",
  sr_location = temp_dir,
  algal_mask = FALSE,
  save = TRUE, 
  save_location = temp_dir
)

# Don't need the object loaded for this example
rm(full_sr)
gc()
```

## 4. Perform WQP-siteSR matchups

We'll now match our downloaded chlorophyll *a* *in-situ* data to the stacked siteSR dataset using a 5-day window on either side of the *in-situ* measurements.

```{r join-wqp-sr}
matchup_path <- match_siteSR_to_WQP(
  wqp_path = file.path(temp_dir, "chla_harmonized.feather"),
  siteSR_path = file.path(temp_dir, "siteSR_DSWE1_full_concatenation.feather"),
  # Grab path from existing object
  site_list_path = siteSR_paths["siteSR_site_list"],
  save_location = file.path(temp_dir, "chla_siteSR_DSWE1_matchups.parquet"),
  # This is also the default time_window
  time_window = "5 days"
)
```

## 5. Apply the Handoff Corrections

Now that *in-situ* and SR matchups are complete, we can use `apply_handoffs()` to harmonize across Landsat missions.

```{r apply-handoffs}
# Apply the Gardner polynomial correction
handoff_path <- apply_handoffs(
  input_path = matchup_path,
  handoff_path = siteSR_paths["handoffs"],
  correction_method = "Gardner_poly",
  sat_target = "LS7",
  algal_mask = FALSE,
  save_location = file.path(temp_dir, "chla_siteSR_DSWE1_matchups.parquet")
)
```

## 6. Access the Results

`apply_handoffs()` has now written the final, harmonized dataset to the path returned by `handoff_path`. We can open it "lazily" with `{arrow}` and bring a preview into memory using `collect()`.

```{r preview}
# Open the new harmonized dataset
harmonized_matchups <- arrow::open_dataset(handoff_path)

# Row and column results
dim(harmonized_matchups)

# Preview the results
harmonized_matchups %>%
  head(10) %>%
  collect()
```

If you are confident that you have sufficient memory to work with the dataset without using `{arrow}` then you could also consider loading it fully into memory before doing further analysis. 
```{r final-to-df}
harmonized_matchups_df <- as.data.frame(harmonized_matchups)
```
