---
title: "matchups"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{matchups}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Once downloading of AquaMatch products is complete, the `{AquaMatchr}` package can be used to match the *in-situ* Water Quality Portal data with the remote sensing data. We provide functions to complete this task over a couple of steps:

1. `build_sr()`: A function that stacks the component files of either siteSR or lakeSR into a single table
2. `match_siteSR_to_WQP()`: A function that joins the *in-situ* WQP data to siteSR data at shared locations within a window of time
3. `apply_handoffs()`: A function that applies intermission handoff coefficients to harmonize data across Landsat missions

We'll walk through the process below, beginning with loading the package:
```{r setup, message = FALSE, warning = FALSE}
library(AquaMatchr)
```

# Stacking SR data

The siteSR and lakeSR datasets are both downloaded as multiple files through their respective functions in this package. This is necessary due to the way they are stored online but also because of the size and complexity of their component parts. Both are available through the Environmental Data Initiative as separate files by DSWE filter (1, confident water vs 1a, confident water and algal mask) and by Landsat mission (4, 5, 7, 8, 9). 

The `AquaMatchr::build_sr()` function can be used to stack files for either siteSR or lakeSR into a single Arrow Table object, which may also be exported to a file. The user must indicate which of the two products is being stacked (`which_sr`), the location where their component files have been stored on disk (`sr_location`), whether the algal mask version is being stacked or not (`algal_mask`), and whether the resulting stack should be saved to disk (`save`). If custom filenames have been given to the SR files then they can be provided using the `sr_files` argument. If `save` is TRUE, then users should provide either a directory or `.feather` file location to export the output to as well via `save_location`.

```{r, eval=FALSE}
# No sr_files provided here, so default names used by download_siteSR()
# are assumed by the function
build_sr(
    which_sr = "siteSR",
    # Mission 4,5,7,8,9 files stored in hypothetical data/ folder
    sr_location = "data",
    algal_mask = FALSE,
    save = TRUE,
    # Exports to .feather file in this folder
    save_location = "data/matchup_output"
)
```

We use Arrow Tables from the [{arrow} R package](https://arrow.apache.org/docs/r/articles/data_objects.html) in order to help save on memory because the files involved are very large when combined. It is possible to convert these to more familiar tibbles or data frames if desired, but please be aware that they will take up more memory and thus be more likely to cause R to crash if you do so. For example, the output of `build_sr()` above can be converted to a data.frame using the code below, but would become much larger.

```{r, eval=FALSE}
test <- build_sr(
    which_sr = "siteSR",
    sr_location = "data",
    algal_mask = FALSE,
    save = TRUE,
    save_location = "data/matchup_output"
)

test_df <- as.data.frame(test)
```

Many data manipulation operations can be completed with Arrow Tables using `dplyr` syntax, so we encourage users to investigate these options. (See, for example, [Data analysis with dplyr syntax](https://arrow.apache.org/docs/r/articles/data_wrangling.html)).

# Joining *in-situ* data with siteSR

As with the previous step, we use the {arrow} package to make joining the siteSR dataset to the *in-situ* WQP possible. This is accomplished through the `match_siteSR_to_WQP()` function. Join operations between two very large datasets can cause memory usage in R to balloon to unmanageable sizes, so we perform the joining process entirely using the {arrow} package. This package doesn't need to pull the entire dataset into your computer's memory in order to analyze it, which is what makes huge joins like this possible. It is also remarkably fast. As a minor tradeoff, the output file must be stored as .parquet file, which can be be opened using {arrow}.

The `match_siteSR_to_WQP()` function requires:

+ `wqp_path`: the path to either a .csv or .feather file containing the *in-situ* dataset
+ `siteSR_path`: the path to the siteSR dataset
+ `site_list_path`: the path to the .csv file containing the siteSR site list (included in downloads done using `download_siteSR()`)
+ `time_window`: the amount of time on either side of the *in-situ* measurements that should be used to match to siteSR overpass times
+ `save_location`: The path where a parquet file containing the output should be saved

```{r, eval=FALSE}
# Example matching harmonized chlorophyll a to siteSR with a 5-day window:
match_siteSR_to_WQP(
  wqp_path = "data/chla_harmonized.feather",
  siteSR_path = "data/matchup_output/siteSR_DSWE1_full_concatenation.feather",
  site_list_path = "data/siteSR_collated_WQP_NWIS_sites_with_NHD_info_2025-06-04.csv",
  save_location = "data/matchup_output/chla_siteSR_matchups.parquet",
  time_window = "5 days"
)
```

Once the function successfully runs and the output is written to a .parquet file, it can be read back into R. Depending on the size of the file and your computer's available memory, it may also be feasible to convert this into a regular data frame. While the join process itself would be very memory intensive if not done outside of memory with {arrow}, the product of `match_siteSR_to_WQP()` should have relatively fewer rows than the siteSR dataset itself. Therefore, once you've read the file back in with `arrow::open_dataset()`, it may be worth seeing if it is manageable to convert it back to a data frame or tibble:

```{r, eval=FALSE}
# Open dataset as an Arrow FileSystemDataset
matchups <- arrow::open_dataset(
  sources = "data/matchup_output/chla_siteSR_matchups.parquet"
)

# May convert to data frame if desired
matchups_df <- as.data.frame(matchups)

# Remove original
rm(matchups)
gc()
```

